Waving and swaying

### 2013-12-13 Paper/grass.tex
update_draft_dec13

\section{Introduction}
Different kinds of motion shown by field canopy and sea grass due to wind and water flow respectively are very common 
phenomena which affects the hydrodynamic conditions. Altered hydrodynamic conditions can influence number of environmental 
variables and processes such as transport of heat, water, vapor, gas , pollens, sediments, nutrients, contaminants, dissolved oxygen 
and other different kind of nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature; 
wind damage to jungle and crop canopies. Driven by such applications number of scientific work regarding flow over canopies has been done over 
last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
of canopy scales.\newline
A key feature of flow through submerged vegetation or terrestrial canopies is the existence of strong shear near top of canopy due to
different amount of drag experienced by fluid in and above the canopy, existence of this strong shear is manifested as
existence of inflection point near top of canopy in velocity profile. A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
etc.\newline
Previous studies of monami and homani have been done under various assumption such as existence of mixing layer profile near top of canopy; Viscosity playing minor role 
in generating cohrent eddies; Irrotationality of flow; Ignorance of drag in stability analysis. While assumption of viscosity playing minor role seems to be 
a good approximation in large Raynols number, assumption of irrotionality may not be good approximation and more importantly exclusion of drag of canopy in stability 
calculation of mean flow may not be right approximation given the fact that it is drag due to canopy which is responsible for existence of inflection point in mean velocity
profile\newline     
In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
medium which is coupled to flow field via a drag term. We did linear stability analysis of flow predicted by model in two different scenario via solving 
modified orr-sommerfield equation, our calculation predicts threshold condition for waving, we we further confirm threshold using CFD calculation as well.

\section{The model}
We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow. Presence of grass modifies the flow of fluid via exerting drag, this drag force is
taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
shape of grass at that location, which can be obtained by assuming grass to be in quasi-equilibrium at any moment and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
\begin{equation}
 \mathbf{f}=-N\mathbf{f_{d}}
\end{equation}
where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density 
Drag force exerted by fluid on grass depends on its density $\rho$, flow velocity u, orientation and diameter of grass, which we model as 
\begin{equation}
 \mathbf{f_{d}}=C_{N}\rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
\end{equation}
 \frac{\del}{\del s}\left(T\hat{t}+N\hat{n}\right) & +\mathbf{f_{d}}+\mathbf{f_{buoy}} = 0\\
 \frac{\del M}{\del s}&-N = 0\\
 M &= B\kappa
where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass and $\kappa$ is curvature 
of grass.  
\begin{center}
\begin{figure}[ht]
\begin{minipage}[b]{5cm}

@@

\centering
 \includegraphics[width=4.0cm,height=4.0cm]{Grass_orientation}
\caption{Typical grass orientation}
\end{minipage}
\hspace{1.0cm}
\begin{minipage}[b]{5cm}
\centering
 \includegraphics[width=5.0cm,height=5.5cm]{Grass_orientation2} 
\caption{Forces and moment on a section of grass}
\end{minipage}
\end{center}
For flexible grass whose bending stiffness is really low, we assume that monami/honami is response to oscillations in flow velocity of fluid. In that case we can safely ignore
set of equation(3).

@@

 In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow 
in presence of grass, we investigate stability of base state obtained through equation of motion. We consider small perturbations $u, v, p$  associated 

@@

In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
of having been produced through equation of motion itself.\newline
We performed stability analysis on two different kind of base flow namely 1) flow driven by constant pressure gradient between two parallel plate  2) constant pressure gradient driven 

@@

 seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get modified orr-sommerfield equation which is an eigenvalue problem
whose solution will provide frequency and wavelength of eddies traveling over canopy.

>>>>

\section{Introduction}
Rich set of dynamics are exhibited by sea grasses due to its interaction with water flow which can considerably affect the hydrodynamic conditions near the 
sea grasses, resulting change in hydrodynamic conditions can influence number of environmental 
variables and processes relative to unobstructed flow such as transport of heat, water, vapor, sediments, contaminants, dissolved oxygen 
and various other nutrients; plant growth and biomass production (known as thigmomorphogenesis); vegetation and land temperature etc. 
%Driven by such applications number of scientific work regarding flow over canopies has been done over 
%last three decades. One of main development is recognition that major contribution to canopy turbulence arises from coherent eddies
%of canopy scales.\newline
Among many dynamics shown by flexible grasses due to their interaction with water flow, the response to steady current in form of large amplitude coherent oscillations known as
Monami (Ackerman and Okubo 1993) is focus of this paper. Similar phenomenon of large amplitude oscillation of terrestrial canopy in wind have also been observed named as Honami 
(Inoue 1993). Experimental observations (Grizzle 1996) have shown that monami is observed when mean flow speed exceed a certain threshold value. 
\newline
A key feature of flow through submerged vegetation is the existence of strong shear near top of canopy due to
different amount of drag experienced by fluid in and above the canopy, existence of strong shear is manifested by
existence of an inflection point near top of canopy in velocity profile. 
%A crucial feature of a velocity inflection is that it induces a hydrodynamic instability
%process which sets the pattern of coherent eddies propagating over terrestrial and submerged canopies. In response to coherent eddies canopies
%may exhibit synchronous large-amplitude waving (termed as monami in aquatic setting by Ackerman and Okubo[1993], and honami by Inoue (1955) in terrestrial setting). 
%Monami (Honami) can be observed when flow velocity is above certain threshold value, which can depends on flow depth, elasticity of canopy material
%etc.\newline
Inspired by work of Raupach (1996) for the case of terrestrial grass, previous studies (Ghisalberti 2002) of monami 
have been done under assumption that a pattern of coherent eddies propagating downstream is set by a hydrodynamic instability of shear layer 
similar to that of free shear flow, response of sea grass to coherent eddies propagating downstream results in large amplitude synchronous oscillations known as monami. 
Among some of drawback of previous theory includes
(1) While atmospheric flow through terrestrial vegetation seems similar to aquatic flow through marine grass, atmospheric flow is unbounded in vertical direction 
but aquatic flow is not so its not obvious that similar mechanism will hold in aquatic case too (2) The proposed hydrodynamic instability mechanism of shear layer produced 
by presence of grass does not take into account the presence of grass itself (3) It does not predict any threshold condition for waving to happen 
\newline     
In this paper we propose a coupled fluid-canopy model to study dynamics of flexible canopy and flow field. Canopies are represented by a continuous 
medium which is coupled to flow field via a drag term. We did linear stability analysis of flow without ignoring presence of grass via solving 
a generalized orr-sommerfield equation with presence of grass. Our calculation predicts threshold a threshold criteria for waving which is characterized by Reynolds number.
We further show that for flexible marine grass monami is caused by instability of flow above the grass rather than by shear layer. We confirm our analysis by comparing 
prediction from theory with existing data

\section{The model}
We propose a fully coupled model in order to comprehend dynamics of canopy and fluid flow, drag force excerted by grasses is
taken as an additional body force in momentum equation of fluid. We approximate presence of grass via continuous field whose shape at any location is approximated as averaged 
shape of grass at that location which can be obtained by assuming grass to be in quasi-equilibrium at any moment.
% and solving force balance and momentum balance for grass  
\begin{equation}
\rho \left(\bu_{t}+\bu.\grad\bu \right) = -\grad P+\mu\grad^{2}\bu +\mathbf{f}+\rho\mathbf{g}
\end{equation}
\begin{equation}
 \mathbf{f}=-N\mathbf{f_{d}}
\end{equation}
where $\mathbf{f}$ is local average drag force per unit volume; $\mathbf{f_{d}}$ is drag force per unit length of grass; N is grass number density. 
Drag force exerted by fluid on grass depends on fluid density $\rho$, flow velocity $\bu$, orientation and diameter of grass which we model as 
\begin{equation}
 \mathbf{f_{d}}=C_{N}\rho\bu_{N}^{2}d\hat{n}+C_{T}\rho\bu_{T}^{2}d\hat{t}
\end{equation}

%\begin{equation}
%\begin{split}
% \frac{\del}{\del s}\left(T\hat{t}+N\hat{n}\right) & +\mathbf{f_{d}}+\mathbf{f_{buoy}} = 0\\
% \frac{\del M}{\del s}&-N = 0\\
% M &= B\kappa
%\end{split}

%\end{equation}
where $C_{N}$ and $C_{T}$ are normal and tangential drag coefficients respectively; $\bu_{T}$, $\bu_{N}$ are velocity vector along and normal to grass.

@@

%\begin{figure}[ht]
%\begin{minipage}[b]{5cm}
%\centering
% \includegraphics[width=4.0cm,height=4.0cm]{Grass_orientation}
%\caption{Typical grass orientation}
%\end{minipage}
%\hspace{1.0cm}
%\begin{minipage}[b]{5cm}
%\centering
% \includegraphics[width=5.0cm,height=5.5cm]{Grass_orientation2} 
%\caption{Forces and moment on a section of grass}
%\end{minipage}
%\end{figure}
%\end{center}
%For flexible grass whose bending stiffness is really low, we assume that monami/honami is response to oscillations in flow velocity of fluid. In that case we can safely ignore
%set of equation(3).

@@

In order to understand frequency and wavelength associated with passage of dominant eddies over grass which are manifestation of instability of mean flow, 
we investigate stability of base flow profile which can be obtained by solving steady state solution of equation(1). We consider small perturbations $u, v, p$  associated 

@@

%In most of recent work on honami/monami base profile have been approximated to a plane mixing layer profile having inflection point near top of canopy, It is also argued that 
%It is instability associated with inflection of mean velocity which is responsible for coherent eddies traveling on top of canopies. In our calculation we get base velocity
%profile by hydrodynamic balance between pressure, drag and diffusive force which also results in a velocity profile with inflection point near top of canopy with additional property 
%of having been produced through equation of motion itself.\newline
We performed stability analysis on two different kind of base flow namely (1) flow driven by constant pressure gradient between two parallel plate (2) constant pressure gradient driven 

@@

which is zero outside the grass. Seeking wave solution in form $\left(u,v,\psi \right)= \left(\hat u, \hat v, \hat\phi \right)e^{ikx+\sigma}$, we get generalized orr-sommerfield equation 
which is an eigenvalue problem whose solution provides frequency and wavelength of eddies traveling over canopy.
